A tool set for random number generation on GPUs in R

### 2022-01-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3090 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{clrng::rnorm}. The best performance we have seen using \fct{clrng::rnorm} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 512 * 128)
R> system.time(clrng::rnorm(c(10000, 10000), streams = streams,
+    Nglobal = c(512, 128), type = "double"))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.045   0.004   0.049
\end{CodeOutput} 
\end{CodeChunk}

@@

<<monthDataForScript, include=FALSE, purl=TRUE>>-
knitr::kable(month)
@

\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 256 * 64)
R> month_gpu <- vclMatrix(month, type = "integer")
R> system.time(result_month <- clrng::fisher.sim(month_gpu, 1000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.320   0.002   0.322
\end{CodeOutput}
\begin{CodeInput}
R> result_month$threshold
\end{CodeInput}
\begin{CodeOutput}
## [1] -47955
\end{CodeOutput}
\begin{CodeInput}
R> result_month$simNum
\end{CodeInput}
\begin{CodeOutput}
## [1] 1015808
\end{CodeOutput}
\begin{CodeInput}
R> result_month$counts
\end{CodeInput}
\begin{CodeOutput}
## [1] 409435
\end{CodeOutput}
\begin{CodeInput}
R> result_month$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.403
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
R> system.time(result_week <- clrng::fisher.sim(week_GPU, 10000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   1.949   0.008   1.960
\end{CodeOutput}
\begin{CodeInput}
R> result_week$threshold
\end{CodeInput}
\begin{CodeOutput}
## [1] -54990
\end{CodeOutput}
\begin{CodeInput}
R> result_week$simNum
\end{CodeInput}
\begin{CodeOutput}
## [1] 10010624
\end{CodeOutput}
\begin{CodeInput}
R> result_week$counts
\end{CodeInput}
\begin{CodeOutput}
## [1] 1351
\end{CodeOutput}
\begin{CodeInput}
R> result_week$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.000135
\end{CodeOutput} 
\end{CodeChunk}

>>>>

We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{clrng::rnorm}. The best performance we have seen using \fct{clrng::rnorm} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(512 * 128)
R> system.time(clrng::rnorm(c(10000, 10000), streams = streams,
+    Nglobal = c(512, 128), type = "double"))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.049   0.000   0.048
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 256 * 64)
R> month_gpu <- vclMatrix(month, type = "integer")
R> system.time(result_month <- clrng::fisher.sim(month_gpu, 1000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.319   0.002   0.320
\end{CodeOutput}
\begin{CodeInput}
R> result_month$threshold
\end{CodeInput}
\begin{CodeOutput}
## [1] -47955
\end{CodeOutput}
\begin{CodeInput}
R> result_month$simNum
\end{CodeInput}
\begin{CodeOutput}
## [1] 1015808
\end{CodeOutput}
\begin{CodeInput}
R> result_month$counts
\end{CodeInput}
\begin{CodeOutput}
## [1] 409435
\end{CodeOutput}
\begin{CodeInput}
R> result_month$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.403
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
R> system.time(result_week <- clrng::fisher.sim(week_GPU, 10000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   1.942   0.007   1.948
\end{CodeOutput}
\begin{CodeInput}
R> result_week$threshold
\end{CodeInput}
\begin{CodeOutput}
## [1] -54990
\end{CodeOutput}
\begin{CodeInput}
R> result_week$simNum
\end{CodeInput}
\begin{CodeOutput}
## [1] 10010624
\end{CodeOutput}
\begin{CodeInput}
R> result_week$counts
\end{CodeInput}
\begin{CodeOutput}
## [1] 1351
\end{CodeOutput}
\begin{CodeInput}
R> result_week$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.000135
\end{CodeOutput} 
\end{CodeChunk}
